---
title: "Unsupervised clustering"
output:
  html_document:
    theme: journal
  pdf_document: default
---

# Library
```{r library}
library(ade4)
library(vegan)
library(gclus)
library(cluster)
library(rgl)
library(scatterplot3d)
library(e1071)
library(multcompView)
library(labdsv)
library(kohonen)
library(MASS)
```

#Functions
##Rand index (Hubert and Arabie) #
```{r AdjRkl}
#Plus ou moins un test du Chi-deux

AdjRkl <- function(part1, part2)
{
  confusion <- table(part1, part2)
  
  n <- sum(confusion)
  a <- sum(choose(confusion[confusion>1], 2))
  b <- apply(confusion, 1, sum)
  b <- sum(choose(b[b>1], 2))
  c <- apply(confusion, 2, sum)
  c <- sum(choose(c[c>1], 2))
  
  Rexp <- b*c/choose(n,2)
  (a - Rexp) / (.5*(b+c) - Rexp )
}
```

##Choose number of clusters
```{r Rousseuw}
Rousseuw <- function(x, dist, dendro = NULL, k) 
{
  asw <- numeric(nrow(x))
  for (k in 2:(nrow(x)-1))
    {
      sil <- silhouette(cutree(dendro, k=k), dist)
      asw[k] <- summary(sil)$avg.width
    }
  k.best <- which.max(asw)
  
  # The plot is produced by function plot.silhouette {cluster}
  plot(1:nrow(x), asw, type="h",
       main="Silhouette-optimal number of clusters",
       xlab="k (number of groups)", ylab="Average silhouette width")
  axis(1, k.best, paste("optimum",k.best,sep="\n"), col="red", font=2,
       col.axis="red")
  points(k.best, max(asw), pch=16, col="red", cex=1.5)
  cat("", "Silhouette-optimal number of clusters k =", k.best, "\n",
      "with an average silhouette width of", max(asw), "\n")
}
```

##Choose number of clusters for PAM
```{r Rousseuw_PAM}
Rousseuw_PAM <- function (x, dist, k)
{
  asw <- numeric(nrow(spe))
  
  for (k in 2:(nrow(spe)-1))
    asw[k] <- pam(spe.ch, k, diss=TRUE)$silinfo$avg.width
  k.best <- which.max(asw)
  cat("", "Silhouette-optimal number of clusters k =", k.best, "\n",
      "with an average silhouette width of", max(asw), "\n")
  plot(1:nrow(spe), asw, type="h", main="Choice of the number of clusters",
       xlab="k (number of groups)", ylab="Average silhouette width")
  axis(1, k.best, paste("optimum",k.best,sep="\n"), col="red", font=2,
       col.axis="red")
  points(k.best, max(asw), pch=16, col="red", cex=1.5)
}
```

##plot cluster with colors on dendrogram
```{r hcoplot}
hcoplot <- function(tree, diss, k, title=paste("Reordered dendrogram from",deparse(tree$call),sep="\n")) 
{
  require(gclus)
  gr <- cutree(tree, k=k)
  tor <- reorder.hclust(tree, diss)
  plot(tor, hang=-1, xlab=paste(length(gr),"sites"), sub=paste(k,"groups"), main=title)
  so <- gr[tor$order]
  gro <- numeric(k)
  for (i in 1:k) {
    gro[i] <- so[1]
    if (i<k) so = so[so!=gro[i]]
  }
  rect.hclust(tor, k=k, border=gro+1, cluster=gr)
  legend("topright", paste("Group",1:k), pch=22, col=2:(k+1), bty="n")
}
```

##Silouhette plot
```{r silouhette_plot}
silouhette_plot <- function(x, dist, dendro, k)
{
  cutg <- cutree(dendro, k=k)
  sil <- silhouette(cutg, dist)
  silo <- sortSilhouette(sil)
  rownames(silo) <- row.names(x)[attr(silo,"iOrd")]
  plot(silo, main="Silhouette plot - Chord",
       cex.names=0.8, col=cutg+1, nmax.lab=100)
}

```
Ce que le graphe nous montre :

* Couleurs : répartition des individus dans un groupe --> 1 couleur = 1 groupe
 
* Axe Silhouette width : mesure de similarité de l'individu avec son propre cluster --> Barre colorée grande (proche de 1) = individu en harmonie avec son cluster

* Nombres à coté des barres colorées : numéro du cluster, nombre d'individus dans le cluster, moyenne des mesures de similarité des individus composant le cluster --> valeur proche de 1 = cluster homogène

* Average silhouette width : homogénéité moyenne --> valeur proche de 1 = clustering homogène possible

##Plot cluster on a map if coordinates are available
```{r cluster_map}
cluster_map <- function(coor, cutree)
{
  plot(coor, asp=1, type="n", main="Four Ward groups",
       xlab="x coordinate (km)", ylab="y coordinate (km)")
  lines(coor, col="light blue")
  text(50, 10, "Upstream", cex=1.2, col="red")
  text(25, 115, "Downstream", cex=1.2, col="red")
  
  # Add the four groups
  grw <- cutree
  k <- length(levels(factor(grw)))
  for (i in 1:k) {
    points(coor[grw==i,1], coor[grw==i,2], pch=i+20, cex=3, col=i+1, bg=i+1)
  }
  text(coor, row.names(coor), cex=0.8, col="white", font=2)
  legend("bottomright", paste("Group",1:k), pch=(1:k)+20, col=2:(k+1),
         pt.bg=2:(k+1), pt.cex=2, bty="n")
}
```

##Abundant species in clusters
```{r}
abundant_groups <- function (x, cluster)
{
  groups <- as.factor(spe.kmeans$cluster)
  spe.means <- matrix(0, ncol(spe), length(levels(groups)))
  row.names(spe.means) <- colnames(spe)
  for(i in 1:ncol(spe)) 
  {
    spe.means[i,] <- tapply(spe[,i], spe.kmeans$cluster, mean)
  }
  return(spe.means)
}
```

##Indicator species in clusters (Dufrene and Legendre)
```{r indicator}
indicator <- function(x, y, k)
{
  das.D1 <- dist(data.frame(das=y[,1], row.names=rownames(y)))
  dasD1.kmeans <- kmeans(das.D1, centers=k, nstart=100)
  dasD1.kmeans$cluster
  
  #Indicator species for this typology of the sites
  iva <- indval(x, dasD1.kmeans$cluster)
  iva
  
  #Table of the significant indicator species
  gr <- iva$maxcls[iva$pval<=0.05]
  iv <- iva$indcls[iva$pval<=0.05]
  pv <- iva$pval[iva$pval<=0.05]
  fr <- apply(x>0, 2, sum)[iva$pval<=0.05]
  fidg <- data.frame(group=gr, indval=iv, pvalue=pv, freq=fr)
  fidg <- fidg[order(fidg$group, -fidg$indval),]
  return(fidg)
}
```

# Data
## Read 
```{r data}
data(doubs)
data(iris)
data(lascaux)
```

## Rename
```{r}
#Doubs
spe <- doubs$fish
env <- doubs$env
spa <- doubs$xy
```

```{r}
#Iris
iris.var <- iris[,c(1:4)]
```

#Matrices de distances
```{r distance_matrix}
# Verify if empty sites exist
apply(spe, 1, max)

# Remove the site where there is no fish
spe = spe[-8,]
env = env[-8,]
spa = spa[-8,]

# Check heterogeneity
boxplot(spe) 

# Normalize (scale between 0 and 1)
# Standardize (scale between -3 and 3)
spe.norm <- decostand (spe, "normalize")
boxplot(spe.norm)

# Matrix distance
spe.ch <- dist(spe.norm)
```


#Dendrograms
```{r dendrogram}
par(mfrow = c(2,3))

spe.ch.single <- hclust(spe.ch, method = "single")
plot(spe.ch.single, main = "Single")

spe.ch.complete <- hclust(spe.ch, method = "complete")
plot(spe.ch.complete, main = "Complete")

spe.ch.ward <- hclust(spe.ch, method = "ward.D")
plot(spe.ch.ward, main = "Ward.D")

spe.ch.average <- hclust(spe.ch, method = "average")
plot(spe.ch.average, main = "Average")

spe.ch.centroid <- hclust(spe.ch, method = "centroid")
plot(spe.ch.centroid, main = "Centroid")

spe.ch.median <- hclust(spe.ch, method = "median")
plot(spe.ch.median, main = "Median")
```

##Which method is better ?
###Cophenetic correlation
```{r cophenetic}
spe.ch.ward.coph <- cophenetic(spe.ch.ward)
cor(spe.ch, spe.ch.ward.coph) 

spe.ch.single.coph <- cophenetic(spe.ch.single)
cor(spe.ch, spe.ch.single.coph) 

spe.ch.complete.coph <- cophenetic(spe.ch.complete)
cor(spe.ch, spe.ch.complete.coph) 

spe.ch.median.coph <- cophenetic(spe.ch.median)
cor(spe.ch, spe.ch.median.coph) 

spe.ch.centroid.coph <- cophenetic(spe.ch.centroid)
cor(spe.ch, spe.ch.centroid.coph) 

spe.ch.average.coph <- cophenetic(spe.ch.average)
cor(spe.ch, spe.ch.average.coph)
```

UPGMA seems to be the best method. But this method don't always give the best result, the best solution

###Shepard-like diagrams
```{r Shepard}
par(mfrow=c(2,2))

plot(spe.ch, spe.ch.single.coph, xlab="Chord distance",
     ylab="Cophenetic distance", asp=1, xlim=c(0,sqrt(2)), ylim=c(0,sqrt(2)),
     main=c("Single linkage",paste("Cophenetic correlation ",
                                   round(cor(spe.ch, spe.ch.single.coph),3))))
abline(0,1)
lines(lowess(spe.ch, spe.ch.single.coph), col="red")

plot(spe.ch, spe.ch.complete.coph, xlab="Chord distance",
     ylab="Cophenetic distance", asp=1, xlim=c(0,sqrt(2)), ylim=c(0,sqrt(2)),
     main=c("Complete linkage", paste("Cophenetic correlation ",
                                      round(cor(spe.ch, spe.ch.complete.coph),3))))
abline(0,1)
lines(lowess(spe.ch, spe.ch.complete.coph), col="red")

plot(spe.ch, spe.ch.average.coph, xlab="Chord distance",
     ylab="Cophenetic distance", asp=1, xlim=c(0,sqrt(2)), ylim=c(0,sqrt(2)),
     main=c("UPGMA", paste("Cophenetic correlation ",
                           round(cor(spe.ch, spe.ch.average.coph),3))))
abline(0,1)
lines(lowess(spe.ch, spe.ch.average.coph), col="red")

plot(spe.ch, spe.ch.ward.coph, xlab="Chord distance",
     ylab="Cophenetic distance", asp=1, xlim=c(0,sqrt(2)), ylim=c(0,max(spe.ch.ward$height)),
     main=c("Ward clustering", paste("Cophenetic correlation ",
                                     round(cor(spe.ch, spe.ch.ward.coph),3))))
abline(0,1)
lines(lowess(spe.ch, spe.ch.ward.coph), col="red")
```

Points have to be grouped
Ward is better according to this method and quasi with cophenetic method

##How many groups?
###Fusion level values 
```{r fusion_levels}
par(mfrow=c(2,2))
summary(spe.ch.single)  #List of available results

plot(spe.ch.single$height, nrow(spe):2, type="S",
     main="Fusion levels - Chord - Single",
     ylab="k (number of clusters)", xlab="h (node height)", col="grey")
text(spe.ch.single$height, nrow(spe):2, nrow(spe):2, col="red", cex=0.8)

plot(spe.ch.complete$height, nrow(spe):2, type="S",
     main="Fusion levels - Chord - Complete",
     ylab="k (number of clusters)", xlab="h (node height)", col="grey")
text(spe.ch.complete$height, nrow(spe):2, nrow(spe):2, col="red", cex=0.8)

plot(spe.ch.average$height, nrow(spe):2, type="S",
     main="Fusion levels - Chord - UPGMA",
     ylab="k (number of clusters)", xlab="h (node height)", col="grey")
text(spe.ch.average$height, nrow(spe):2, nrow(spe):2, col="red", cex=0.8)

plot(spe.ch.ward$height, nrow(spe):2, type="S",
     main="Fusion levels - Chord - Ward",
     ylab="k (number of clusters)", xlab="h (node height)", col="grey")
text(spe.ch.ward$height, nrow(spe):2, nrow(spe):2, col="red", cex=0.8)
```

h (node heigth) --> distance between groups
The distance must be high between groups to become the good method
See all the check-methods to decide
Goal : the more parcimonious groupment with homogeneous groups

###Rousseeuw quality index
```{r Rousseuw index}
Rousseuw(x = spe, dist = spe.ch, dendro = spe.ch.ward, k = 4)
```

Plot average silhouette widths (using Ward clustering) for all partitions
except for the trivial partition in a single group (k=1).
First, create an empty vector in which the asw values will be written
good cluster number = high silhouette value

Two clusters are the best choice but the second better choice is 4. 
It is the better choice in ecology

Choose 4 groups according to all methods
```{r}
k <- 4
```

```{r cutree}
#Cut the tree
spebc.single.g <- cutree(spe.ch.single, k)
spebc.complete.g <- cutree(spe.ch.complete, k)
spebc.average.g <- cutree(spe.ch.average, k)
spebc.ward.g <- cutree(spe.ch.ward, k)

table(spebc.average.g, spebc.ward.g) #To see common groups in both methods
table(spebc.average.g, spebc.single.g) 
table(spebc.average.g, spebc.complete.g) 
table(spebc.single.g, spebc.complete.g)
table(spebc.single.g, spebc.ward.g)
table(spebc.ward.g, spebc.complete.g)
```

Agree for the majority but differencies.

##How to quantify differencies?
```{r}
AdjRkl(spebc.single.g, spebc.complete.g) 
AdjRkl(spebc.single.g, spebc.average.g) 
AdjRkl(spebc.single.g, spebc.ward.g)
AdjRkl(spebc.ward.g, spebc.complete.g)
AdjRkl(spebc.ward.g, spebc.average.g) 
AdjRkl(spebc.average.g, spebc.complete.g)
```

The more is the index, the more the methods are confusive
Ward and Complete give very similar results

##Final partition
```{r final_partition}
silouhette_plot(spe, spe.ch, spe.ch.ward, 4)
```

Higher is the coefficient, better is
High coefficient : homogeneous group
Low coefficient : no homogeneous but they have to be placed somewhere


##Dendrogram with colors of groups
```{r dendro_color}
hcoplot(spe.ch.ward, spe.ch, k)
```


##Plot of the 4 clusters on a map of the Doubs river
```{r map_doubs}
cluster_map(spa, spebc.ward.g)
```

There is a continuity but with some problems
Ecology can answer these problems : river cut off with a barrier, etc...

#AGNES
Agglomerative hierarchical clustering

```{r agnes_ward}
#Use Ward method according to check-methods above
agnes1 <- agnes(spe.norm, method = "ward")
par(mfrow = c(1,2))
agnes1
plot(agnes1)
```

Very good agglomerative coefficient close to 1
2 well differenciated cluster (blank between)
But we can differenciate other groups (littlest bar)

```{r agnes_UPGMA}
#Use method UPGM average - The best according to cophenetic correlation #
agnes2 <- agnes(spe.norm, method = "average")
par(mfrow = c(1,2))
agnes2
plot(agnes2)
```

Lower agglomerative coefficient

##AGNES VS Ward Method
```{r agnesVSward}
#Choose the number of cluster
clas.agnes <- cutree(agnes1,k)
table(cutree(agnes1,k))
table(cutree(agnes1,k), spebc.ward.g)
AdjRkl(cutree(agnes1,k), spebc.ward.g)
```

Perfect confusive table - Coefficient = 1 - Same results

```{r clusplot_agnes_ward}
clusplot(spe.norm, clas.agnes, 
         color = T, shade = T, labels = 2, lines = 1,
         main = "Clusplot AGNES")
clusplot(spe.norm, spebc.ward.g, 
         color = T, shade = T, labels = 2, lines = 1,
         main = "Clusplot Ward")
```

If two groups are linked by a line, there are different
If not, there are not different
Groups with lower within ss are well differenciated on the graph

```{r}
table(spebc.ward.g, clas.agnes)
```

Same graphics - Same results

#DIANA
Divisive hierarchical clustering

```{r diana_euclidian}
diana1 <- diana(spe.norm, metric = "euclidian")
par(mfrow = c(1,2))
diana1
plot(diana1)
```

Coefficient lower than agnes cause diana group down.
It is easier to aggregate than divide

```{r diana_manhattan}
diana1 <- diana(spe.norm, metric = "manhattan")
par(mfrow = c(1,2))
diana1
plot(diana1)
```


##DIANA VS AGNES & Ward Method

```{r dianaVSward}
clas.diana <- cutree(diana1,k)
table(clas.diana)
table(spebc.ward.g, clas.diana) #DIANA VS Ward 
table(clas.agnes, clas.diana) #DIANA VS AGNES
```

Same results because Ward and AGNES give the same results
Aggree with the majority but some differencies

```{r}
clusplot(spe.norm, clas.diana, color = T, shade = T, labels = 2, lines = 1,
         main = "Clusplot DIANA")
```

```{r}
AdjRkl(clas.diana, spebc.ward.g)
```

```{r}
#In the same table
clas.all <- cbind(spebc.ward.g, clas.agnes, clas.diana)
clas.all
```

#kmeans
Minimize the sum of squares from points to cluster centres
Very sensitive to the initial centres, choose a nstart > 1
nstart -> avoid biaised center position by choosen many times

```{r kmeans}
spe.kmeans <- kmeans(spe.norm, centers = 4, nstart = 100)
spe.kmeans
```

cluster sum of squares -> closer to 0 is an homogeneous group

```{r kmeans_objects}
#Different objects
spe.kmeans$cluster
spe.kmeans$centers
spe.kmeans$totss
spe.kmeans$withinss
spe.kmeans$tot.withinss
spe.kmeans$betweenss #totss - tot.withinss
spe.kmeans$iter
```


##kmeans VS Ward

```{r kmeansVSward}
table(spe.kmeans$cluster, spebc.ward.g)
#Agree except with one

cbind(spe.kmeans$cluster, spebc.ward.g)
#See which individuals are not classified in the same group

AdjRkl(spebc.ward.g, spe.kmeans$cluster)
#Very close to 1 --> Method agree

clusplot(spe.norm, spe.kmeans$cluster, color = T, shade = T, labels = 2, lines = 1)
#Same graph as Ward method
```



##Plot of the 4 k-means clusters on a map of the Doubs river

```{r}
cluster_map(spa, spe.kmeans$cluster)
```


## CascadeKM - Kmeans partitioning
Wrapper of the kmeans function. Creates several partitions
forming a cascade from a small to a large number of groups

```{r KM}
spe.KM.cascade <- cascadeKM(spe.norm, inf.gr = 2, sup.gr = 10, iter = 100, criterion = "ssi")
par(mfrow = c(1,1))
plot(spe.KM.cascade, sortg = T)
```

If you want two groups -> horizontal line from two groups (y axis) 
and see how many objects are classed into two groups
If you want four groups -> same method
Choose the number of cluster where the ssi are higher --> well diferenciated groups
Here : 3 but is it really better than 4 groups? Need knowledge in ecology

##ratio between sum of squares
```{r KM_sumsquares}
summary(spe.KM.cascade)
names(spe.KM.cascade)
spe.KM.cascade$results
ssi.sse <- spe.KM.cascade$results[2,]/spe.KM.cascade$results[1,]*100
barplot(ssi.sse) 
#Ratio must be lowest as possible
```

##Repartition of sites
```{r KM_repartition}
spe.KM.cascade$partition
```


#Partitioning Around Medoids - PAM
A more robust version of kmeans

```{r PAM}
spe.pam <- pam(spe.norm, 4, diss = F, metric = "euclidian", stand = F)
par(mfrow=c(1,2))
plot(spe.pam)
spe.pam
```

##PAM VS kmeans & Ward method

```{r PAMvskmeansvsward}
table(spe.pam$clustering, spe.kmeans$cluster)
AdjRkl(spe.pam$clustering, spe.kmeans$cluster)
AdjRkl(spe.pam$clustering, spebc.ward.g)

clusplot(spe.norm, spe.pam$clustering, color = T, shade = T, labels = 2, lines = 1)
#All groups are well differenciated with PAM
```

##Choice of the number of clusters with PAM

```{r}
Rousseuw_PAM(spe, spe.ch, k = k)
```

3 or 4 groups --> 4 is often the best choice in ecology

## Plot of the 4 PAM clusters on a map of the Doubs river
```{r}
cluster_map(spa, spe.pam$cluster)
```


# Crossed with the environment table
##Prepare data kmeans

```{r}
par(mfrow = c(2,2))
boxplot(sqrt(env$alt) ~ spe.kmeans$cluster, 
        main = "Altitude", las = 1, ylab = "sqrt(alt)", col = 2:5, varwidth = T)
boxplot(log(env$slo) ~ spe.kmeans$cluster, 
        main = "Slope", las = 1, ylab = "sqrt(slo)", col = 2:5, varwidth = T)
boxplot(env$oxy ~ spe.kmeans$cluster, 
        main = "Oxygen", las = 1, ylab = "sqrt(oxy)", col = 2:5, varwidth = T)
boxplot(sqrt(env$amm) ~ spe.kmeans$cluster, 
        main = "Ammonium", las = 1, ylab = "sqrt(amm)", col = 2:5, varwidth = T)
```


##Prepare data PAM
```{r}
boxplot(sqrt(env$alt) ~ spe.pam$clustering, 
        main = "Altitude", las = 1, ylab = "sqrt(alt)", col = 2:5, varwidth = T)
boxplot(log(env$slo) ~ spe.pam$clustering, 
        main = "Slope", las = 1, ylab = "sqrt(slo)", col = 2:5, varwidth = T)
boxplot(env$oxy ~ spe.pam$clustering, 
        main = "Oxygen", las = 1, ylab = "sqrt(oxy)", col = 2:5, varwidth = T)
boxplot(sqrt(env$amm) ~ spe.pam$clustering, 
        main = "Ammonium", las = 1, ylab = "sqrt(amm)", col = 2:5, varwidth = T)
```

Visualize some gradients

##Normality test
```{r normality}
shapiro.test(resid(lm(sqrt(env$alt) ~ as.factor(spe.kmeans$cluster)))) #NS
shapiro.test(resid(lm(log(env$slo) ~ as.factor(spe.kmeans$cluster)))) #NS
shapiro.test(resid(lm(env$oxy ~ as.factor(spe.kmeans$cluster)))) #NS
shapiro.test(resid(lm(sqrt(env$amm) ~ as.factor(spe.kmeans$cluster)))) #NS
```

##Homoscedasticity
```{r homoscedasticity}
bartlett.test(sqrt(env$alt), as.factor(spe.kmeans$cluster)) #Non
bartlett.test(log(env$slo), as.factor(spe.kmeans$cluster)) #NS
bartlett.test(env$oxy, as.factor(spe.kmeans$cluster)) #NS
bartlett.test(sqrt(env$amm), as.factor(spe.kmeans$cluster)) #NS
```

##Which variable explain the clusters - Test 
```{r test}
summary(aov(log(env$slo) ~ as.factor(spe.kmeans$cluster))) #Oui
summary(aov(env$oxy ~ as.factor(spe.kmeans$cluster))) #Oui
summary(aov(sqrt(env$amm) ~ as.factor(spe.kmeans$cluster))) #Oui
kruskal.test(sqrt(env$alt) ~ as.factor(spe.kmeans$cluster)) #Oui
```

##Graphics
```{r multibox}
datslo <- as.data.frame(cbind(log(env$slo), spe.kmeans$cluster))
names(datslo) <- c("slo", "kmean.cl")
datslo
multcompBoxplot(slo ~ kmean.cl, data = datslo)
#Two groups differenciated

datalt <- as.data.frame(cbind(sqrt(env$alt), spe.kmeans$cluster))
names(datalt) <- c("alt", "kmean.cl")
datalt
multcompBoxplot(alt ~ kmean.cl, data = datalt)
#Two or three groups well differenciated

datoxy <- as.data.frame(cbind(env$oxy, spe.kmeans$cluster))
names(datoxy) <- c("oxy", "kmean.cl")
datoxy
multcompBoxplot(oxy ~ kmean.cl, data = datoxy)
#Three groups well differenciated

datamm <- as.data.frame(cbind(sqrt(env$amm), spe.kmeans$cluster))
names(datamm) <- c("amm", "kmean.cl")
datamm
multcompBoxplot(amm ~ kmean.cl, data = datamm)
#Three groups well differenciated
```

Environment represent the river course
Downstream and upstream caracteristics differenciate groups

##Kmeans abundance
Species that represent each groups are species 
with a frecuency higher than the mean

```{r}
abundant_groups(spe, spe.kmeans$cluster)
```

###Mean species abundances of the four groups
```{r}
group1 <- round(sort(abundant_groups(spe, spe.kmeans$cluster)[,1], decreasing=TRUE), 2)
group2 <- round(sort(abundant_groups(spe, spe.kmeans$cluster)[,2], decreasing=TRUE), 2)
group3 <- round(sort(abundant_groups(spe, spe.kmeans$cluster)[,3], decreasing=TRUE), 2)
group4 <- round(sort(abundant_groups(spe, spe.kmeans$cluster)[,4], decreasing=TRUE), 2)
```

###Species with abundances greater than group mean species abundance

```{r}
group1.domin <- which(group1>mean(group1))
group1
group1.domin
#7 species

group2.domin <- which(group2>mean(group2))
group2
group2.domin
#12 species - Redundant with group 1

group3.domin <- which(group3>mean(group1))
group3
group3.domin
#5 species 

group4.domin <- which(group4>mean(group1))
group4
group4.domin
#22 species
```


##Species indicator values (Dufrene and Legendre)
Divide the sites into 4 groups depending on the distance to the source of the river

```{r}
indicator(spe, env, k)
```

##Export the result to a CSV file (to be opened in a spreadsheet)
```{r csv_indicator}
write.csv(indicator(spe, env, k), "IndVal-das.csv")
```

#Fanny
Fuzzy Analysis Clustering

##iris
```{r fanny_iris}
fanny1 <- fanny(iris.var, k = 3, diss = F)
fanny1
par(mfrow = c(1,2))
plot(fanny1)
```
One group well differenciated
More difficult for the others

###Plot membership
```{r}
par(mfrow = c(1,1))
plot(c(1:150), fanny1$membership[,1], type = "l", col = "blue")
lines(c(1:150), fanny1$membership[,2], type = "l", col = "red")
lines(c(1:150), fanny1$membership[,3], type = "l", col = "green")
```
See the probability of an individual to be in a cluster

```{r magicbox}
s3d <- scatterplot3d(fanny1$membership, color = fanny1$cluster, type = "h", angle = 100, scale.y = 0.7, pch = 16, main = "Pertinence")
plot3d(fanny1$membership, col = fanny1$cluster)
```
3D representations

##Doubs
```{r fanny_doub}
fanny2 <- fanny(spe.norm, k = 4, diss = F)
fanny2
par(mfrow = c(1,2))
plot(fanny2)

par(mfrow = c(1,1))
plot(c(1:29), fanny2$membership[,1], type = "l", col = "blue")
lines(c(1:29), fanny2$membership[,2], type = "l", col = "red")
lines(c(1:29), fanny2$membership[,3], type = "l", col = "green")
lines(c(1:29), fanny2$membership[,4], type = "l", col = "black")

plot3d(fanny2$membership, col = fanny2$cluster, type = "s")
```


###Plot of the 4 Fanny clusters on a map of the Doubs rive
```{r}
cluster_map(spa, fanny2$cluster)
```

#Cmeans
Fuzzy version of kmeans

##Iris
```{r cmeans_iris}
cmeans1 <- cmeans(iris.var, 3 , 50)
cmeans1
cmeans1$membership

plot3d(cmeans1$membership, col = cmeans1$cluster, type = "p")

plot(iris, col = cmeans1$cluster)
plot(iris, col = iris$Species)
```

Compare the cluster with the real species

##Doubs

```{r cmeans_doubs}
cmeans2 <- cmeans(spe.norm, 4 , 50)
cmeans2
cmeans2$membership

plot3d(cmeans2$membership, col = cmeans2$cluster, type = "s")
```

###Plot of the 4 cmeans clusters on a map of the Doubs rive
```{r}
cluster_map(spa, cmeans2$cluster)
```

#Clara
##Prepare data

```{r clara_morpho}
morpho <- lascaux$morpho

clara1 <- clara(lascaux$morpho, k = 5)
clara1
par(mfrow = c(1,2))
plot(clara1)

table(clara1$clustering)
```


#SOM
##Iris
SOM Tools Box better with Matlab - But R can do it too
But it's ugly - maybe#

somgrid = size of the map, normally at least 5*sqrt(n)

```{r SOM_iris}
set.seed(700) #the random number always start from this point
sqrt(150)*5 #150 individuals in Iris
somiris <- som(as.matrix(iris[,1:4]), grid = somgrid(7,9,"hexagonal")) #or rectangular for less neighbors
names(somiris)
```

###Graphics map
```{r SOM_map}
par(mfrow = c(1,1))
par(mfrow = c(2,2))
plot(somiris, type = "codes")
#All flower types possible according to the data

plot(somiris, type = "counts")
#How many individuals in each flower type

labiris <- c(rep("s", 50), rep("e", 50), rep ("i", 50))
plot(somiris, type = "mapping", labels = labiris)
#Which species in each flower type

plot(somiris, type = "mapping", pch = rep(1:3, rep (50,3)), col = rep(1:3, rep (50, 3)))
#Other representation of the species
```

###Plot a dendrogram with Ward method
```{r}
par(mfrow = c(1,1))
h1 <-  hclust(dist(somiris$codes), method = "ward.D")
plot(h1)

plot(somiris, "mapping", pch = rep(1:3, rep (50,3)), 
     col = rep(1:3, rep (50, 3)), bgcol = cutree(h1,3)+4)
#Bgcol = background color --> Represent the groups determined by hclust

som.predict <- predict(somiris, trainY = factor(iris[,5]))
table(iris[,5], som.predict$prediction)
AdjRkl(iris[,5], som.predict$prediction)
#The method is very close to the real species
```

##Doubs 
```{r}
set.seed(700)
somdoubs <- som(as.matrix(spe.norm), grid = somgrid(4,7,"hexagonal"))

plot(somdoubs, type = "codes")
plot(somdoubs, type = "counts")
plot(somdoubs, type = "quality")
plot(somdoubs, type = "mapping", labels = factor(rep(1:29,1)))

par(mfrow = c(1,1))
hdoubs <-  hclust(dist(somdoubs$codes), method = "ward.D")
plot(hdoubs)
plot(somdoubs, "mapping", labels = factor(rep(1:29,1)), bgcol = cutree(hdoubs,4)+4)
```

###SOM vs All

```{r SOMvskmeans}
som.predict.doubs.k <- predict(somdoubs, trainY = spe.kmeans$cluster)
table(spe.kmeans$cluster, som.predict.doubs.k$prediction)
AdjRkl(spe.kmeans$cluster, som.predict.doubs.k$prediction)
```

```{r SOMvsagnes}
som.predict.doubs.a <- predict(somdoubs, trainY = clas.agnes)
table(clas.agnes, som.predict.doubs.a$prediction)
AdjRkl(clas.agnes, som.predict.doubs.a$prediction)
```

```{r SOMvsdiana}
som.predict.doubs.d <- predict(somdoubs, trainY = clas.diana)
table(clas.diana, som.predict.doubs.d$prediction)
AdjRkl(clas.diana, som.predict.doubs.d$prediction) 
```

```{r SOMvsPAM}
som.predict.doubs.d <- predict(somdoubs, trainY = spe.pam$cluster)
table(spe.pam$cluster, som.predict.doubs.d$prediction)
AdjRkl(spe.pam$cluster, som.predict.doubs.d$prediction)
```

```{r SOMvscmeans}
som.predict.doubs.d <- predict(somdoubs, trainY = cmeans2$cluster)
table(cmeans2$cluster, som.predict.doubs.d$prediction)
AdjRkl(cmeans2$cluster, som.predict.doubs.d$prediction)
```

